### 2022-10-06 submission/manuscript.Rmd
Bring in Table 1 from Singletons papper

---
output:
  word_document: default
  pdf_document:
    keep_tex: true
geometry: margin=1.0in
font-size: 11pt
header-includes:
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left]{lineno}
  - \linenumbers
  - \raggedright
editor_options:
  chunk_output_type: console
bibliography: references.bib
csl: asm.cls
---

```{r, echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(here))

opts_chunk$set(echo = FALSE)

inline_hook <- function(x){

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps){
			# treat as integer
			formatted <- format(x, digits=0, big.mark = ",")

		} else {
			# treat as floating point number
			formatted <- format(x, digits=2, nsmall=2, big.mark = ",")
		}
	} else {
		formatted <-x
	}

  formatted

@@

knit_hooks$set(inline=inline_hook)

format_pct <- function(x, my_digits=1){

	format(x, digits=my_digits, nsmall=my_digits)

}
```

@@

@@

***Choice of datasets.*** My analysis incorporated community distribution data from 12 studies that represented bacterial communities from diverse environments. The specific studies were selected because their data was publicly accessible through the Sequence Read Archive, the original investigators sequenced the V4 region of the 16S rRNA gene using paired 250 nt reads, and my previous familiarity with the data. The use of paired 250 nt reads to sequence the V4 region resulted in a near complete two-fold overap of the V4 region resulting in high quality contigs with a low sequencing error rate. These data were processed through the standard sequence curation pipeline to generate operational taxonomic units (OTUs) using the mothur software package. The original studies generated the sequence data by pooling separate PCR products that were generated by amplifying the V4 region from the bacterial DNA in each sample. Because pooling equimolar quantities of DNA is frought with difficulties, it was common to observe wide variation in the number of sequences in each sample (Figure S1, Table 1).

@@

***Evidence that not rarefying data can impact results in original studies.*** To assess the impact of using alternatives to rarefaction, I reassessed hypotheses from the human, mice, and peromyscus studies. These studies were selected because they originated from my research group and know the datasets well. First, the study that published the human dataset was interested in the difference in the fecal microbial communities of patients with and without colorectal cancer. Here, I divided the patients into peope with and without screen relevant neoplasia (SRN). Among the alpha diversity metrics, the test were non-significant with similar effect sizes; however, the richness estimates obtained using parametric appraoches detected statistically significant differences in richness. Although the tests using Bray-Curtis distances were significant regardless of whether rarefaction was used, the effect sizes were smaller for the distances calculated using the CLR, VST, CSS transformations.

@@

### Re-analysis of previously published datasets
* Human study: Did not alter effect size or significance of alpha or beta diversity, but did result in reduced effect sizes for measures of richness and non-parametric estimators of richness; breakaway detected a difference

@@

<div id="refs"></div>

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

>>>>

---
bibliography: references.bib
output:
  pdf_document:
    keep_tex: true
csl: asm.csl
geometry: margin=1.0in
header-includes:
 - \usepackage{booktabs}
 - \usepackage{longtable}
 - \usepackage{array}
 - \usepackage{multirow}
 - \usepackage{wrapfig}
 - \usepackage{float}
 - \usepackage{colortbl}
 - \usepackage{pdflscape}
 - \usepackage{tabu}
 - \usepackage{threeparttable}
 - \usepackage{threeparttablex}
 - \usepackage[normalem]{ulem}
 - \usepackage{makecell}
 - \usepackage{setspace}
 - \doublespacing
 - \usepackage[left]{lineno}
 - \linenumbers
 - \modulolinenumbers
 - \usepackage{helvet} % Helvetica font
 - \renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
 - \usepackage[T1]{fontenc}
 - \usepackage[shortcuts]{extdash}
---

```{r, echo=FALSE}
options(tidyverse.quiet = TRUE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(glue))

opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("message" = FALSE)
opts_chunk$set("cache" = FALSE)

inline_hook <- function(x, digits=2){

  if(is.list(x)){
    x <- unlist(x)
  }
  if(is.numeric(x)){
      paste(format(x,big.mark=',', digits=digits, nsmall=digits, scientific=FALSE))
  } else {
      paste(x)
  }
}
knitr::knit_hooks$set(inline=inline_hook)

package_version <- function(package){

  paste(unlist(packageVersion(package)), collapse='.')

@@

oxford_comma <- function(x, digits=2) {

  x <- map_chr(x, inline_hook, digits=digits)

  if(length(x) < 2){
    x
  } else if(length(x) == 2){
    paste(x, collapse = " and ")
  } else {
    paste(paste(x[-length(x)], collapse=", "), x[length(x)], sep=", and ")
  }
}
```

@@

```{r datasets}
table_1 <- read_tsv(here('data/process/study_summary_statistics.tsv'))

n_sample_range <- table_1 %>%
  pull(n_samples) %>%
  range() %>% as.integer()

n_med_sequence_range <- table_1 %>%
  pull(median) %>%
  range() %>% as.integer()

low_fold_samples <- table_1 %>%
  filter(fold_difference < 2) %>%
  pull(directory)

high_fold_range <- table_1 %>%
  filter(fold_difference >= 2) %>%
  pull(fold_difference) %>%
  range() %>%
  round(digits=1)
```

@@

***Choice of datasets.*** I selected 16S rRNA gene sequence data from from 12 studies that characterized the variation in bacterial communities from diverse environments (Table 1). The specific studies were selected because their data was publicly accessible through the Sequence Read Archive, the original investigators sequenced the V4 region of the 16S rRNA gene using paired 250 nt reads, and my previous familiarity with the data. The use of paired 250 nt reads to sequence the V4 region resulted in a near complete two-fold overap of the V4 region resulting in high quality contigs with a low sequencing error rate [@Kozich2013]. These data were processed through the standard sequence curation pipeline to generate operational taxonomic units (OTUs) using the mothur software package [@Kozich2013; @Schloss2009]. The original studies generated the sequence data by pooling separate PCR products that were generated by amplifying the V4 region of the 16S rRNA gene from the bacterial DNA in multiple samples. Because pooling equimolar quantities of DNA is frought with difficulties, it was common to observe wide variation in the number of sequences in each sample (Figure S1).

@@

***Evidence that not rarefying data can impact results in original studies.*** To assess the impact of using alternatives to rarefaction, I reassessed hypotheses from the human, mice, and peromyscus studies. These studies were selected because they originated from my research group and I know the datasets well. First, the study that published the human dataset was interested in the difference in the fecal microbial communities of patients with and without colorectal cancer. Here, I divided the patients into peope with and without screen relevant neoplasia (SRN). Among the alpha diversity metrics, the test were non-significant with similar effect sizes; however, the richness estimates obtained using parametric appraoches detected statistically significant differences in richness. Although the tests using Bray-Curtis distances were significant regardless of whether rarefaction was used, the effect sizes were smaller for the distances calculated using the CLR, VST, CSS transformations.

* Human study: Did not alter effect size or significance of alpha or beta diversity, but did result in reduced effect sizes for measures of richness and non-parametric estimators of richness; breakaway detected a difference

@@

@@

<div id="refs"></div>
\bibliography{ref}
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

**Table 1. Summary of studies used in the analysis.** For all studies, the number of sequences used from each study was rarefied to the smallest sample size. A graphical represenation of the distribution of sample sizes for each study and the samples that were removed from each study are provided in Figure S1.

\small

```{r results_f}
table_1 %>%
  arrange(nice_name) %>%
  mutate(
    nice_name = glue("{nice_name}\\ {reference}"),
    total_seqs=format(total_seqs, big.mark=",", trim=TRUE),
    median=format(as.integer(median), big.mark=",", trim=TRUE),
    min=format(min, big.mark=",", trim=TRUE),
    max=format(max, big.mark=",", trim=TRUE),
    range = glue("{min}\\Hyphdash*{max}"),
    n_samples = n_samples) %>%
  select(nice_name, n_samples, total_seqs, median, range, sra_study) %>%
  kableExtra::kable(format="markdown", booktabs=TRUE, escape=F, align="lrrrrr",
    linesep="",
    col.names = linebreak(
      c(
        "\\textbf{Study\\nobreakspace{}(Ref)}",
        "\\textbf{Samples}",
        "\\textbf{Total}\n\\textbf{sequences}",
        "\\textbf{Median}\n\\textbf{sequences}",
        "\\textbf{Range of}\n\\textbf{sequences}",
        "\\textbf{SRA study}\n\\textbf{accession}"
      ),
      align="c")
  )
```
\normalsize
